# Missing values

```{r}
library(readr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(patchwork)
```

## Cancer Risk Dataset

```{r}
plot_missing <- function(df, percent=FALSE) {
  missing_patterns <- data.frame(is.na(df)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup()
  
  pure_pattern <- missing_patterns[,-which(names(missing_patterns) == "count")]
  pure_pattern$incomplete <- as.logical(rowSums(pure_pattern))
  missing_patterns$incomplete <- pure_pattern$incomplete

  tidymissing <- missing_patterns %>% 
    rownames_to_column() %>%
    gather(key, value, -c(count, rowname, incomplete))
  
  tidymissing$rowname <- as.factor(as.numeric(tidymissing$rowname))

  main_plot <- ggplot(tidymissing, aes(x=fct_reorder(key,-value*count, sum),y=fct_rev(rowname),fill=value, alpha=incomplete)) +
    geom_tile(color="white") + theme_classic() +
    scale_fill_discrete(type=c("grey", "purple")) +
    scale_alpha_manual(values = c(1,0.5)) +
    annotate("text", x = length(unique(tidymissing$key))/2+0.5, y = which(rev(pure_pattern$incomplete)==FALSE), label="complete cases") + 
    xlab('variable') +
    ylab('missing pattern') + 
    theme(legend.position = "none")
  
  missing_by_pattern <- missing_patterns %>%
  select(count, incomplete) %>%
  rownames_to_column() 
  
  missing_by_pattern$rowname <- as.factor(as.numeric(missing_by_pattern$rowname))
  missing_by_pattern$percentage <- missing_by_pattern$count/sum(missing_by_pattern$count)*100
  
  missing_by_pattern$plotby <- missing_by_pattern$count 
  y_lab_pattern = "row count"
  if (percent==TRUE) {
    missing_by_pattern$plotby <- missing_by_pattern$percentage
    y_lab_pattern = "% rows"
  }

  pattern_bar <- ggplot(missing_by_pattern, aes(x=fct_rev(rowname), y=plotby, alpha=incomplete)) + 
    geom_bar(stat="identity",fill="blue") + 
    scale_alpha_manual(values = c(1,0.5)) +
    coord_flip() +
    ylab(y_lab_pattern) +
    theme(panel.background = element_rect(fill = NA), axis.title.y = element_blank(), panel.grid.major.y = element_blank(), 
          panel.grid.major.x = element_line(colour="grey"), panel.grid.minor.x = element_line(colour="grey"), panel.grid.minor.y = element_blank(),
          panel.border = element_rect(fill = NA), legend.position = "none")
  
  temp_df = apply(data.frame(is.na(df)),2,sum)
  ylabel = "num rows\n missing:"
  if (percent==TRUE) {
    temp_df = apply(data.frame(is.na(df)),2,mean)*100
    ylabel = "% rows\n missing:"
  }
  
  missing_by_col <- data.frame(temp_df) %>% 
    rownames_to_column() %>%
    filter(rowname!='incomplete')

  colnames(missing_by_col) <- c('key', 'pattern_count')

  col_bar <- ggplot(missing_by_col, aes(x=fct_reorder(key, -pattern_count), y = pattern_count))+geom_bar(stat="identity",fill="lightblue") +
    ylab(ylabel) +
    theme(panel.background = element_rect(fill = NA), axis.title.x = element_blank(), panel.grid.major.x = element_blank(), 
          panel.grid.major.y = element_line(colour="grey"), panel.grid.minor.x = element_blank(),panel.border = element_rect(fill = NA)) +
    ggtitle("Missing value patterns")
  
  if (percent==TRUE) {
    pattern_bar <- pattern_bar + ylim(0, 100)
    col_bar <- col_bar + ylim(0, 100)
  }
  
  col_bar + (ggplot()+theme_void())+ main_plot + pattern_bar + plot_layout(ncol=2, widths=c(4,1), heights=c(1,3))
}
```

The original risk data frame does not have any missing values since GBD Results Tool has already ignored all the observations with missing value. 

```{r}
names(cancer_risk)<-abbreviate(names(cancer_risk), minlength = 4)
plot_missing(cancer_risk)
```

So I get another smaller dataset to get another perspective of missing value and the dataset is as follows:

```{r}
cancerdata <- read_csv("cancerdata.csv")
head(cancerdata)
```

Our original dataset is downloaded from the source https://ghdx.healthdata.org/. It shows the information about in the global scale in 2019, the death rates of each kind of cancer caused by different external risks for all ages. We can see that the dataset is not in a tidy format. There are many useless variables that only contains one value. So below we transformed the data into a tidy format using only the variables of our interests.

```{r}
newdata <- cancerdata[,c("cause", "rei", "val")]
names(newdata) <- c("type", "risk", "death_rate")

tidydata <- newdata %>%
  pivot_wider(names_from = risk, values_from = death_rate) %>%
  column_to_rownames(var="type")
head(tidydata)
```

After transformation, we get a new data frame. Each observation is the death rates of one kind of cancer caused by different external risks. There is a lot of missing values in our dataset because not all types of cancer are related to all kinds of risks. For example, some types of cancer may be induced by air pollution, but will not be induced by alcohol use. 

In order to see the relationship between cancer types and external risks, we analyze the missing values in this data frame below.

```{r}
names(tidydata)<-abbreviate(names(tidydata), minlength = 6)
plot_missing(tidydata)
```

From the missing pattern plot we can see that for each of the risks air pollution, drug use, other environmental risks, unsafe sex and low physical activity, their values are missing for all other types of cancer except one. That means theses risks are highly correlated to one specific kind of cancer and people need to be care of those risks if they would like to avoid the specific types of cancer. 

We can also see that if data is missing for risk air pollution, then it is likely to be missing in for other environmental risks. It makes sense since air pollution is also one kind of environmental risk. It is chosen to be a separated variable because it accounts for a large proportion of environmental risks.

Another thing of our notice is that risk tobacco is almost not missing for all kinds of cancer, which means it can be the induction of the majority types of cancer. So in order to protect ourselves from being attacked by cancer, we should try to avoid use tobacco a lot.

We do not have a complete case here because no type of cancer would be induced by all the external risks.

Below, by explicitly drawing missing values in each observation, we can have a better understanding of similarities in different kinds of cancer.

```{r}
tidycancer <- tidydata %>%
  rownames_to_column("type") %>%
  gather(key, value, -type) %>%
  mutate(missing = ifelse(is.na(value), 1, 0)) %>%
  group_by(key) %>%
  mutate(Std = ifelse(value-mean(value, na.rm = TRUE)==0, 0, (value-mean(value, na.rm = TRUE))/sd(value, na.rm=TRUE))) %>% 
  ungroup()

ggplot(tidycancer, aes(x=fct_reorder(key, -missing, sum), y=fct_reorder(type, -missing, sum), fill=Std)) +
  geom_tile(color="white") +
  scale_fill_gradient2(low="blue", mid="white", high = "yellow", na.value = "black") + 
  xlab("type of cancer") +
  ylab("risk") +
  theme_bw()
```

From the plot above, tracheal, bronchus, and lung cancer is affected by external risks the most since it can be induced by majority of the risks listed. It also has the highest death rate because we can see many of the blocks are yellow which means the rate is very high.

Uterine cancer, thyroid cancer and gallbladder and biliary tract cancer have the same missing pattern. Colon and rectum cancer and breast cancer have same missing pattern. It suggests that there might be some similarity in these kinds of cancer, though not obvious now, but may help people find treatment to these types of cancer.

Nasopharynx cancer and larynx cancer have same missing pattern. Other pharynx cancer and lip and oral cavity cancer have same missing pattern. And these two missing patterns are also similar to each other. It makes sense since they are all related to throat.


## Cancer by Country Dataset

The dataset mainly has null value for health cost per capita and health cost among GDP.Most of the missing data are from 1990 to 1995.Besides,there are a  few missing data of GDP and GNI. Actually, we lack of the GDP and GNI data about Brazil. Therefore, for our further analysis, when we need GDP an GNI data, we plan to exclude Brazil. When analysing health cost related data,we plan to exclude 1990  and 1995 becuase too concentrated missing data can negatively affect our analysis and there seems no reliable predictive data to fill these NAs.

```{r}
data = read_csv("./data/finalData.csv")
names(data)<-abbreviate(names(data), minlength = 7)
plot_missing(data)
```
